### 2019-01-11 Metrics1.tex
Metrics1

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Econometrics 1. Problem Set 1.
\end{center}
\section*{Problem 1}
 Suppose your calculate estimates for $\beta_0$ and $\beta_1$ by finding the solution to the following minimization problem:
 \begin{align*}
 \underset{b_0, b_1}{\min}\ \sum_{i=1}^n \text{exp} \left\{(y_i - b_0 - b_1x_i)^2\right\}
 \end{align*}
 Write down first-order conditions for the estimates.

 \textbf{Solution}

 \begin{align*}
 \frac{\partial}{\partial b_0}: -2 \sum_{i=1}^n (y_i - b_0 - b_ix_i)e^{(y_i - b_0 - b_ix_i)^2} = 0\\
 \frac{\partial}{\partial b_1}: -2 \sum_{i=1}^n x_i(y_i - b_0 - b_ix_i)e^{(y_i - b_0 - b_ix_i)^2} = 0
 \end{align*}
 \section*{Problem 2}
 In the simple linear regression model $y = \beta_0 + \beta_1x + u$, suppose that $\expect (u) \neq 0$. Letting $a_0 = \expect(u)$, show that the model can always be rewritten with the same slope, but a new
 intercept and error, where the new error has a zero expected value.

 \textbf{Solution}
 Assume $u' = a_0 - u$. Then $\expect(u') = a_0 - a_0 = 0$ and moreover
 \begin{align*}
 &y = \beta_0 + \beta_1 x + a_0 - u' = \beta_0' + \beta_1 x + u''\\
 &\text{where}\ \beta_0' = \beta_0 + a_0\\
 &\ \ \ u'' = -u'\\
 &\text{and}\ \ \expect(u'') = -\expect(u') = 0
 \end{align*}
 Q.E.D.
 \section*{Problem 3}
 Consider the standard simple linear regression model $y = \beta_0 + \beta_1 x + u$
 When $n = 3$, is it possible that the data point with maximal value of $y$ is located below the
 OLS regression line? If answer is yes, provide an example, if no, provide a proof.

 \textbf{Solution}

 Yes, it is possible. Assume $X = (1, 2, 5)$ and corresponded $Y = (2, 3, 4)$ then
 \begin{align*}
 b_1 = \frac{12}{26}\\
 b_0 = \frac{23}{13}
 \end{align*}
 and
 \begin{align*}
 \frac{12}{26} \cdot 5 + \frac{23}{13} = \frac{53}{13} \approx 4.08 > 4
 \end{align*}
 which means that the regression line is lying above the point (5, 4).

 \section*{Problem 4}
  Consider the following relation:
\begin{align*}
y = \sqrt{x} + u
\end{align*}
 where $x$ is uniformly distributed on segment $[0; 1]$ and error term $u$ has zero mean
 conditional on $x$. A random sample of size $n$ is collected: $\left\{x_i, y_i\right\}_{i=1}^n$. Some econometrician
 performs OLS estimation based on a simple linear model 
 \begin{align*}
	y = \beta_0 + \beta_1x + e
\end{align*}
 Based on her estimates $\hat{\beta_0}, \hat{\beta_1}$, the econometrician constructs the predictor for the conditional mean of the dependent variable, i.e. for $\expect(y|x)$, as $\widehat{\expect(y|x)} = \hat{\beta_0} + \hat{\beta_1} x$.
 \begin{enumerate}[a.]
 \item Calculate the expected value of y.
 \item Find the expected values of $\hat{\beta_0}, \hat{\beta_1}$ conditional on realized values of $x$.
 \end{enumerate}

\textbf{Solution}
\begin{enumerate}[a.]
	\item 
	\begin{align*}
	\expect[\sqrt{x}] = \int_{0}^1 \sqrt{x}dx = \frac{2}{3}\\
	\expect[u] = \expect[\expect[u|X]] = \expect[0] = 0\\
	\expect[y] = \frac{2}{3} + 0 = \frac{2}{3}
	\end{align*}
	\item
	\begin{align*}
	\expect[\hat{\beta_1}|x_1, \dots, x_n] = \expect \left[\frac{\sum_{i=1}^n (x_i - \bar{x})(y_i - \bar{y})}{\sum_{i=1}^n (x_i - \bar{x})^2}\bigg|x_1, \dots, x_n\right] =\\
	= \expect \left[\frac{\sum_{i=1}^n (x_i - \bar{x})(\sqrt{x_i} + u_i - \overline{\sqrt{x}} -  \bar{u})}{\sum_{i=1}^n (x_i - \bar{x})^2}\bigg|x_1, \dots, x_n\right]
	\end{align*}
	\end{enumerate}
\end{document}