### 2019-01-18 micro3_2.tex
Micro

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Microeconomics 3. Problem Set 2.
\end{center}
\section*{Problem 1}
Suppose that as a function of factor prices $w = (w_1, w_2) >> 0$, the firms' cost for producing
one unit of output are as follows:
\begin{align*}
c_1(w) = \left(\frac{w_1}{3/4}\right)^{3/4}\left(\frac{w_2}{1/4}\right)^{1/4}\\
c_1(w) = \left(\frac{w_1}{1/4}\right)^{1/4}\left(\frac{w_2}{3/4}\right)^{3/4}
\end{align*}
Let $p_1 = 4$ and $p_2 = \frac{4}{\sqrt{3}}$ be the unit price of the firms' outputs, respectively. Furthermore, assume that the total endowment of the two factors are given by $\bar{z_1} = 1 + \frac{1}{3\sqrt{3}}$ and $\bar{z_2} = 1+\sqrt{3}$. Find the factor prices and allocation in an \textit{interior} equilibrium.

\textbf{Solution}

Since the solution should be interior that means that $c_1(w) = p_1, c_2(w) = p_2$. Hence, 
\begin{align*}
\frac{c_1(w)}{c_2(w)} = \frac{p_1}{p_2}\\
\left(\frac{w_1}{w_2}\right)^{1/2} = \sqrt{3}\\
w_1 = 3w_2\\
\end{align*}
Plugging it into $c_1(w) = p_1$ one can obtain:
\begin{align*}
4w_2 = 4 \to w_2 = 1, w_1 = 3\\
\end{align*}
Now let us find demand vectors for factors:
\begin{align*}
a_1(w) = \begin{pmatrix}
\frac{\partial c_1(w)}{\partial w_1}\\
\frac{\partial c_1(w)}{\partial w_2}
\end{pmatrix} = \begin{pmatrix}
\left(\frac{3w_2}{w_1}\right)^{1/4}\\
\left(\frac{w_1}{3w_2}\right)^{3/4}
\end{pmatrix} = \begin{pmatrix}
1\\
1
\end{pmatrix}\\
a_2(w) = \begin{pmatrix}
\frac{\partial c_2(w)}{\partial w_1}\\
\frac{\partial c_2(w)}{\partial w_2}
\end{pmatrix} = \begin{pmatrix}
\left(\frac{w_2}{3w_1}\right)^{3/4}\\
\left(\frac{3w_1}{w_2}\right)^{1/4}
\end{pmatrix} = \begin{pmatrix}
\dfrac{1}{\sqrt{27}}\\
\\
\sqrt{3}
\end{pmatrix}\\
\end{align*}
Since
\begin{align*}
\frac{a_{11}}{a_{21}} = 1 > \frac{\bar{z_1}}{\bar{z_2}} = \frac{1+3\sqrt{3}}{3\sqrt{3}(1 + \sqrt{3})} = \frac{8 - 2\sqrt{3}}{6\sqrt{3}} > \frac{a_{12}}{a_{22}} = \frac{1}{9}
\end{align*}
that means that solution is indeed interior.
To find the optimal output level $(q_1^*, q_2^*)$ we need the market clearing condition to be satisfied, i.e. we need to solve the following linear system:
\begin{align*}
q_1 \begin{pmatrix}
1\\
1
\end{pmatrix} + q_2 \begin{pmatrix}
\frac{1}{\sqrt{27}}\\
\sqrt{3}
\end{pmatrix} = \begin{pmatrix}
1 + \frac{1}{3\sqrt{3}}\\
1 + \sqrt{3}
\end{pmatrix}
\end{align*}
and the solution is obviously $q_2^* = 1, q_1^* = 1$. Thus, prices for factors are $w = (3, 1)$ and the equilibrium factors' allocation is:
\begin{align*}
\begin{pmatrix}
1\\
1
\end{pmatrix}, \begin{pmatrix}
\frac{1}{3\sqrt{3}}\\
\sqrt{3}
\end{pmatrix}
\end{align*}

\section*{Problem 2}
Suppose that as a function of factor prices $w = (w_1, w_2) >> 0$, the firms' cost for producing
one unit of output are as follows:
\begin{align*}
c_1(w) = \left(\frac{w_1}{2/3}\right)^{2/3}\left(\frac{w_2}{1/3}\right)^{1/3}\\
c_2(w) = \left(\frac{w_1}{1/2}\right)^{1/2}\left(\frac{w_2}{1/2}\right)^{1/2}
\end{align*}
Let
\begin{align*}
p_1 = \frac{6}{2^{\frac{2}{3}}}\text{ and } p_2 = 3
\end{align*}
be the unit price of the firms' outputs, respectively. Furthermore, assume that the total
endowment of the two factors are given by $\bar{z_1} = 2$ and $\bar{z_2} = 1$. Find the factor prices and
allocation in an equilibrium in which firm 2 shuts down. Verify that the firms' decisions are
truly optimal given the factor prices you determined.

\textbf{Solution}

As in previous exercise:
\begin{align*}
\frac{c_1(w)}{c_2(w)} = \frac{p_1}{p_2}\\ \frac{6}{2^{\frac{2}{3}}}\left(\frac{w_1}{w_2}\right)^{1/6} = \frac{6}{2^{\frac{2}{3}}3}\\
w_2 = 3^6w_1
\end{align*}
Now let us find the demand for factors:
\begin{align*}
a_1(w) = \begin{pmatrix}
\frac{\partial c_1(w)}{\partial w_1}\\
\frac{\partial c_1(w)}{\partial w_2}
\end{pmatrix} = \begin{pmatrix}
\left(\frac{2w_2}{w_1}\right)^{1/3}\\
\left(\frac{w_1}{2w_2}\right)^{2/3}
\end{pmatrix} = \begin{pmatrix}
2^{\frac{1}{3}}9\\
\frac{1}{2^{\frac{2}{3}}81}
\end{pmatrix}\\
a_2(w) = \begin{pmatrix}
\frac{\partial c_2(w)}{\partial w_1}\\
\frac{\partial c_2(w)}{\partial w_2}
\end{pmatrix} = \begin{pmatrix}
\left(\frac{w_2}{w_1}\right)^{1/2}\\
\left(\frac{w_1}{w_2}\right)^{1/2}
\end{pmatrix} = \begin{pmatrix}
27\\
\\
\frac{1}{27}
\end{pmatrix}\\
\frac{a_{11}}{a_{21}} = 2\cdot 3^6 > \frac{a_{12}}{a_{22}} = 3^6 > \frac{\bar{z_1}}{\bar{z_2}} = 2
\end{align*}
That means that in the equilibrium firm 1 should shut down.

\section*{Problem 3}
Suppose that for $w = (w_1, w_2) >> 0$, \begin{align*}
a_{11} = \frac{2}{3w_1}\left(\frac{w_1}{2/3}\right)^{2/3}\left(\frac{w_2}{1/3}\right)^{1/3}\\
a_{21} = \frac{1}{3w_2}\left(\frac{w_1}{2/3}\right)^{2/3}\left(\frac{w_2}{1/3}\right)^{1/3}\\
a_{12} = \frac{1}{3w_1}\left(\frac{w_1}{1/3}\right)^{1/3}\left(\frac{w_2}{2/3}\right)^{2/3}\\
a_{22} = \frac{2}{3w_2}\left(\frac{w_1}{1/3}\right)^{1/3}\left(\frac{w_2}{2/3}\right)^{2/3}\\
\end{align*}
where $a_j(w) := (a_{1j}(w); a_{2j}(w))$ is the conditional factor demand of firm $j$ that corresponds
to one unit of output. Assume $p_1 = p_2 = \frac{1}{2}$. Find the equilibrium factor allocation and
prices as a function of factor endowments $\bar{z_1}; \bar{z_2} > 0$. (Be sure to cover boundary equilibria,
if any.)

\textbf{Solution}

Assume $c_1(w), c_2(w)$ are cost functions of firms 1 and 2 for producing 1 unit of output. Then
\begin{align*}
\frac{\partial c_1(w)}{\partial w_1} = a_{11} = \frac{2}{3w_1}\left(\frac{w_1}{2/3}\right)^{2/3}\left(\frac{w_2}{1/3}\right)^{1/3} \to c_1(w) =\left(\frac{w_1}{2/3}\right)^{2/3}\left(\frac{w_2}{1/3}\right)^{1/3} + \varphi(w_2)\\
\end{align*}
where $\varphi(w_2)$ is an arbitrary function of $w_2$.
\begin{align*}
\frac{\partial c_1(w)}{\partial w_2} = \frac{1}{3w_2}\left(\frac{w_1}{2/3}\right)^{2/3}\left(\frac{w_2}{1/3}\right)^{1/3} + \frac{d\varphi(w_2)}{dw_2} = a_{21} \to \varphi(w_2) = const
\end{align*}
if there are no fixed costs then of course $\varphi(w_2) = 0$. Thus, 
\begin{align*}
c_1(w) = \left(\frac{w_1}{2/3}\right)^{2/3}\left(\frac{w_2}{1/3}\right)^{1/3}
\end{align*}
Similarly, 
\begin{align*}
c_2(w) = \left(\frac{w_1}{1/3}\right)^{1/3}\left(\frac{w_2}{2/3}\right)^{2/3}
\end{align*}
Suppose firstly that the equilibrium is interior. Hence
\begin{align*}
\frac{c_1(w)}{c_2(w)} &= \frac{p_1}{p_2}\\
\left(\frac{w_1}{w_2}\right)^{1/3} = 1 &\to w_1 = w_2\\
c_1(w) = p_1 &\to w_1^* = w_2^* = \frac{1}{2^{1/3}3}
\end{align*}
Now let us derive the inferior solution condition.
\begin{align*}
\frac{a_{11}}{a_{21}} &= 2\\
\frac{a_{12}}{a_{22}} &= \frac{1}{2}\\
\end{align*}
\begin{align}\label{eq1}
2 > \frac{\bar{z_1}}{\bar{z_2}} > \frac{1}{2} &\to \frac{\bar{z_2}}{2} < \bar{z_1} < 2\bar{z_2}
\end{align}
Market clearing conditions:
\begin{align*}
q_1 a_1(w) + q_2 a_2(w) = \bar{z}\\
q_1 \begin{pmatrix}
2^{1/3}\\
\frac{1}{2^{2/3}}
\end{pmatrix} + q_2 \begin{pmatrix}
\frac{1}{2^{2/3}}\\
2^{1/3}
\end{pmatrix} = \begin{pmatrix}
\bar{z_1}\\
\bar{z_2}
\end{pmatrix}\\
q_2^* = \frac{2\bar{z_2} - \bar{z_1}}{4 - \frac{1}{2^{2/3}}}, q_1^* = \frac{\bar{z_1}}{2^{1/3}} - \frac{2\bar{z_2} - \bar{z_1}}{8 - 2^{1/3}}
\end{align*}
One can easily check that under \eqref{eq1}, $q_1^* > 0, q_2^* > 0$ i.e. solution is indeed interior. Thus, equilibrium allocation is:
\begin{align*}
\begin{pmatrix}
\bar{z_1} - \frac{2\bar{z_2} - \bar{z_1}}{\frac{8}{2^{1/3}} - 1}\\
\\
\frac{\bar{z_1}}{2^{2/3}} - \frac{2\bar{z_2} - \bar{z_1}}{2^{2/3}8 - 2}
\end{pmatrix}, \begin{pmatrix}
\frac{2\bar{z_2} - \bar{z_1}}{\frac{8}{2^{1/3}} - 1}\\
\\
2^{1/3} \cdot \frac{2\bar{z_2} - \bar{z_1}}{4 - \frac{1}{2^{2/3}}}
\end{pmatrix}
\end{align*}
Assume now that $\frac{\bar{z_1}}{\bar{z_2}} > 2$ that means that the second firm shuts down. Hence
\begin{align*}
\begin{cases}
c_1(w) = p_1\\
\frac{a_{11}}{a_{21}} = \frac{\bar{z_1}}{\bar{z_2}}
\end{cases} \to \frac{2w_2}{w_1} = \frac{\bar{z_1}}{\bar{z_2}} \to w_1 = \frac{1}{3}\left(\frac{\bar{z_2}}{\bar{z_1}}\right)^{1/3}, w_2 = \frac{1}{6}\left(\frac{\bar{z_1}}{\bar{z_2}}\right)^{2/3}
\end{align*}
and the equilibrium allocation is
\begin{align*}
\begin{pmatrix}
\bar{z_1}\\
\bar{z_2}
\end{pmatrix}, \begin{pmatrix}
0\\
0
\end{pmatrix}
\end{align*}
Assume now that $\frac{\bar{z_1}}{\bar{z_2}} < \frac{1}{2}$ that means that the first firm shuts down. Hence
\begin{align*}
\begin{cases}
c_2(w) = p_2\\
\frac{a_{12}}{a_{22}} = \frac{\bar{z_1}}{\bar{z_2}}
\end{cases} \to \frac{w_2}{2w_1} = \frac{\bar{z_1}}{\bar{z_2}} \to w_1 = \frac{1}{6}\left(\frac{\bar{z_2}}{\bar{z_1}}\right)^{2/3}, w_2 = \frac{1}{3}\left(\frac{\bar{z_1}}{\bar{z_2}}\right)^{1/3}
\end{align*}
and the equilibrium allocation is
\begin{align*}
\begin{pmatrix}
0\\
0
\end{pmatrix},
\begin{pmatrix}
\bar{z_1}\\
\bar{z_2}
\end{pmatrix}
\end{align*}
\end{document}