### 2019-01-19 metrics2.tex
Metrics

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Econometrics 1. Problem Set 2.
\end{center}
\section*{Problem 1}
Suppose that the population model determining $y$ is
\begin{align*}
y = \beta_0 + \beta_1 x_1 + \beta_2x_2 + \beta_3x_3 + u
\end{align*}
and this model satisfies the Gauss-Markov assumptions. However, we estimate the model that
omits $x_3$. Let $\tilde{\beta_1}, \tilde{\beta_2}$ be the OLS estimators from the regression of $y$ on $x_1$ and $x_2$. Show
that the expected value of $\tilde{\beta_1}$ (given the values of the independent variables in the sample) is
\begin{align*}
\expect \tilde{\beta_1} = \beta_1 + \beta_3 \frac{\sum_{i=1}^n \hat{r_{i1}x_{i3}}}{\sum_{i=1}^n \hat{r_{i1}}^2}
\end{align*}
where the $\hat{r_{i1}}$ are the OLS residuals from the regression of $x_1$ on $x_2$.

\textbf{Solution}
Plugging $y_i = \beta_0 + \beta_1x_{i1} + \beta_2x_{i2}+\beta_3x_{i3}$ into
\begin{align*}
\tilde{\beta_1} = \frac{\sum_{i=1}^n \hat{r_{i1}}y_i}{\sum_{i=1}^n \hat{r_{i1}}^2}
\end{align*}
one can obtain:
\begin{align} \label{eq1}
\tilde{\beta_1} = \frac{\sum_{i=1}^n \hat{r_{i1}}(\beta_0 + \beta_1x_{i1} + \beta_2x_{i2}+\beta_3x_{i3} + u_i)}{\sum_{i=1}^n \hat{r_{i1}}^2} = \nonumber \\
=\frac{\beta_0 \sum_{i=1}^n \hat{r_{i1}} + \beta_1 \sum_{i=1}^n \hat{r_{i1}}x_{i1} + \beta_2 \sum_{i=1}^n \hat{r_{i1}}x_{i2} + \beta_2 \sum_{i=1}^n \hat{r_{i1}}x_{i2} + \sum_{i=1}^n \hat{r_{i1}}u_i}{\sum_{i=1}^n \hat{r_{i1}}^2}
\end{align}
By properties of OLS estimators (corollary from FOC): $\sum_{i=1}^n \hat{r_{i1}} = 0, \sum_{i=1}^n \hat{r_{i1}}x_{i2} = 0$. Moreover,
\begin{align*}
\sum_{i=1}^n \hat{r_{i1}}x_{i1} = \sum_{i=1}^n \hat{r_{i1}} (\hat{\alpha_0} + \hat{\alpha_1}x_{i2} + \hat{r_{i1}}) = 0 + 0 + \sum_{i=1}^n \hat{r_{i1}}^2
\end{align*}
where $\hat{\alpha_0}, \hat{\alpha_1}$ are OLS estimates for coefficients of regression of $x_1$ on $x_2$. Plugging it into (\ref{eq1}):
\begin{align*}
\tilde{\beta_1} = \beta_1 + \beta_3\frac{\sum_{i=1}^n\hat{r_{i1}}x_{i3}}{\sum_{i=1}^n \hat{r_{i1}}^2} + \frac{\sum_{i=1}^n \hat{r_{i1}}u_i}{\sum_{i=1}^n \hat{r_{i1}}^2}\\
\expect \tilde{\beta_1} = \beta_1 + \beta_3\frac{\sum_{i=1}^n\hat{r_{i1}}x_{i3}}{\sum_{i=1}^n \hat{r_{i1}}^2} + \expect\left[ \frac{\sum_{i=1}^n \hat{r_{i1}}u_i}{\sum_{i=1}^n \hat{r_{i1}}^2} \right]
\end{align*}
\begin{align*}
\hat{r_{i1}} = x_{i1} - \hat{x_{i1}} = x_{i1} - \hat{\alpha_0} - \hat{\alpha_1}x_{i2}\ \to \expect[f(\hat{r_{i1}})|x_{i1}, x_{i2}] = f(\hat{r_{i1}}) \ \forall\ f(\cdot)
\end{align*}
Then,
\begin{align*}
\expect\left[\frac{\sum_{i=1}^n \hat{r_{i1}}u_i}{\sum_{i=1}^n \hat{r_{i1}}^2} \right] = \expect\left[\expect\left[\frac{\sum_{i=1}^n \hat{r_{i1}}u_i}{\sum_{i=1}^n \hat{r_{i1}}^2}\bigg|x_{1}, x_{2}\right] \right] \underset{\text{i.i.d.}}{\propto} \expect\left[\sum_{i=1}^n\hat{r_{i1}}\expect [u_i|x_{i1}, x_{i2}]\right] \underset{\text{G.-M. assump.}}{=} 0 
\end{align*}
Thus, 
\begin{align*}
\expect \tilde{\beta_1} = \beta_1 + \beta_3\frac{\sum_{i=1}^n\hat{r_{i1}}x_{i3}}{\sum_{i=1}^n \hat{r_{i1}}^2}
\end{align*}
\end{document}