### 2019-01-22 macro3_1.tex
Macro

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Macroeconomics 3. Problem Set 1.
\end{center}
\section*{Problem 1}
Consider the following model
\begin{align}\label{eq1}
\underset{\left\{k(t+1), c(t)\right\}_{t=0}^{\infty}}{\max}\ \sum_{t=0}^{\infty} \beta^t u(c(t))\nonumber\\
s.t.\ k(t+1) = k(t) - c(t) \\
c(t) \ge 0, k(t) \ge 0, k(0) > 0\ \text{is given}\nonumber
\end{align}
\begin{enumerate}
	\item Rewrite it recursively. Replace $\beta$ with $\frac{1}{1+\rho}$
	and derive the asset pricing equations for the whole
	cake and for the marginal piece of cake. Explain what are the “dividends” and the “capital gains”
	in each case.
	\item Does the value of the marginal piece of cake increase or decrease over time? Use the asset pricing
	equation and explain the intuition.
	\item Does the value of the whole cake increase or decrease over time? Explain the intuition.
	\item At which rate does the “current value” of the marginal piece of cake change over time? What
	about the “present value”? Use the optimality conditions and explain the intuition.
	\item Assume CRRA utility function,
	\begin{align*}
	u(c(t)) = \frac{c(t)^{1 - \sigma}}{1 - \sigma}, \sigma > 0
	\end{align*}
	Derive the Euler equation. Explain the intuition of how the consumption path depends on the
	intertemporal elasticity of substitution.

	\item Now assume that the cake goes bad as the time goes on, that is the constraint \eqref{eq1} changes to
	\begin{align*}
	k (t + 1) = (1 - \delta) k (t) - c (t)
	\end{align*}
	where $0 < \delta < 1$. Explain how does it change your answers to the previous questions.
	\item Assume $\sigma = 1$ and $\delta = 0$. Find the analytical solutions for the value and policy functions.
\end{enumerate}

\textbf{Solution}

\begin{enumerate}
	\item Let $V(k)$ be the value of the agent who acts optimally and possesses $k$ units of capital now. Then Bellman equation is:
	\begin{align*}
	V(k) = \underset{0 \le c \le k}{\max}\left\{u(c) + \beta V(k - c)\right\} \\
	V(k) = \underset{0 \le c \le k}{\max}\left\{u(c) + \frac{1}{1+\rho} V(k - c)\right\} 
	\end{align*}
	Assume $c^*$ is the optimizer, then
	\begin{align*}
	V(k) = u(c^*) + \frac{1}{1+\rho}V(k-c^*)\\
	\rho = \frac{V(k - c^*) - V(k) + u(c^*)}{V(k) - u(c^*)}
	\end{align*}
	or alternatively:
	\begin{align*}
	\rho = \frac{u(c^*(t)) + V(k^*(t+1)) - V(k^*(t))}{V(k^*(t)) - u(c^*(t))}
	\end{align*}
	(* means optimal trajectory).
	Here $V(k(t+1)) - V(k(t))$ are gross "capital gains" and $u(c^*)$ are gross "dividends".

	By the envelope theorem
	\begin{align*}
	V'(k) = \frac{1}{1+\rho}V'(k-c^*)\\
	\rho = \frac{V'(k-c^*) - V'(k)}{V'(k)}
	\end{align*}
	or alternatively:
	\begin{align*}
	\rho = \frac{V'(k(t+1)) - V'(k(t))}{V'(k(t))}
	\end{align*}
	as one can observe there are no marginal dividends (because there are no production) only $\frac{V'(k(t+1)) - V'(k(t))}{V'(k(t))}$ which describes percentage change of shadow price of capital i.e. marginal "capital gains". 
	\item Of course in general the answer to this question depends on the particular type of function $u()$, however if we assume "good" increasing and concave utility function then price of marginal piece of cake increases over time. It can be inferred both from the asset pricing equation and from the Euler's equation. From the asset pricing equation it is obviously that LHS is nonnegative (its meaning is the interest rate) hence the RHS should be also nonnegative. The denominator of the RHS is nonnegative because economically it is a shadow price of capital or matimatically it is equal to corresponding lagrange multiplier which cannot be negative by the Kuhn and Tucker theorem. Thus, numerator should be also nonnegative, so $V'(k(t+1)) \ge V'(k(t))$ i.e. price of marginal piece of cake increases over time. From the Euler's equation (which is obtained in 5.) $\frac{u'(c(t))}{u'(c(t+1))} = \frac{\lambda(t)}{\lambda(t+1)} = \beta$. Since $\lambda(t) = V'(k(t))$ then $\frac{V'(k(t))}{V'(k(t+1))} = \beta < 1$. That means that $V'(k(t))$ is increasing over time. Intuitivelly it can be understood as follows: the amount of capital is shrinking hence over time the agent values each marginal unit of capital more and more. 
	\item Under some assumptions on the utility function it can be proved that $V(k)$ is strictly increasing. In our problem $k$ is decreasing over time that means that $V(k)$ should also decrease over time, i.e. the value of the whole cake decreases. The amount of cake is decreasing over time, hence less cake $\to$ less it costs.
	\item Using the same equations as in 2. $\frac{V'(k(t))}{V'(k(t+1))} = \beta$, that means that the present value of the marginal piece of cake remains constant: \begin{align*}
	V'(k(t)) = \frac{1}{\beta}V'(k(t-1)) &= \frac{1}{\beta^2}V'(k(t-2)) = \dots = \frac{1}{\beta^t}V'(k(0))\\
	\end{align*}
	But the present value of $V'(k(t))$ is $\beta^t V'(k(t))$ hence 
	\begin{align*}
	\forall\ t\ \beta^t V'(k(t)) = V'(k(0))
	\end{align*}
	At the same time the current value is increasing, which can be inferred from $\frac{V'(k(t))}{V'(k(t+1))} = \beta < 1$.
	\item Assume 
	\begin{align*}
	u(c(t)) = \frac{c(t)^{1 - \sigma}}{1 - \sigma}, \sigma > 0
\end{align*}
then
\begin{align*}
&L = \sum_{t=0}^{\infty} \beta^t(u(c(t)) + \lambda(t)(k(t) - c(t) - k(t+1)))\\
&\begin{cases}\frac{\partial}{\partial c(t)}: u'(c(t)) - \lambda(t) = 0\\
\frac{\partial}{\partial k(t+1)}: -\beta^t \lambda(t) + \beta^{t+1}\lambda(t+1) = 0\\
\end{cases}\\
&\lambda(t) = \beta \lambda(t+1)\\
&u'(c(t+1)) = \lambda(t+1)\\
&\frac{u'(c(t))}{u'(c(t+1))} = \beta\\
\end{align*}
Thus, the Euler's equation is:
\begin{align*}
\left(\frac{c(t+1)}{c(t)}\right)^{\sigma} = \beta\\
\frac{c(t+1)}{c(t)} = \beta^{\frac{1}{\sigma}}
\end{align*}
intertemporal elasticity of substitution is $\frac{1}{\sigma}$. It shows the willingness of agents to substitute consumption over time i.e. to refuse today's consumption in order to increase it tomorrow. If $\frac{1}{\sigma} > 1$ then consumption path is steep: for the sake of increase consumtion today the agent is able to cut his consumption tomorrow, otherwise the consumption path is flat: the agent is not willing to cut tomorrow's consumption hence he should accept lower level of today's consumption (compare to the case $\frac{1}{\sigma} > 1$). See Fig. \ref{fig3}.
\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{plotdraft}
	\caption{}\label{fig3}
\end{figure}
\item Now FOCs are:
\begin{align*}
\begin{cases}\frac{\partial}{\partial c(t)}: u'(c(t)) - \lambda(t) = 0\\
\frac{\partial}{\partial k(t+1)}: -\beta^t \lambda(t) + \beta^{t+1}(1 - \delta)\lambda(t+1) = 0\\
\end{cases}\\
\frac{\lambda(t+1)}{\lambda(t)} = \frac{1}{\beta(1 - \delta)} > \frac{1}{\beta}
\end{align*}
that means that in this case even present value of marginal piece of cake will increase over time:
\begin{align*}
V'(k(t)) = \frac{1}{(1 - \delta)^t}V'(k(0))
\end{align*}
and the rate of increasing of current value $\frac{1}{\beta(1-\delta)}$ goes up compare to the case $\delta = 0$.
\item Assume $\sigma = 1, \delta = 0$ then
from Euler's equation:
\begin{align*}
c(t+1) = \beta c(t)
\end{align*}
it remains to find $c(0)$,
\begin{align*}
&k(t) = k(t+1) + c(t) = k(t+2) + c(t+1) + c(t) = k(t+2) + \beta c(t) + c(t) = \\
&=k(t+3) + c(t+2) + \beta c(t) + c(t) = k(t+3) + \beta^2 c(t) + \beta c(t) + c(t) = \dots =\\
&=k(t + T) + c(t) \sum_{\tau = 0}^{T-1} \beta^{\tau} \\
&\text{since the amount of cake cannot be negative hence for optimal consumption path } k\\
&\text{ should tend to 0 at infinity, }\\
&k(t + T) + c(t) \sum_{\tau = 0}^{T-1} \beta^{\tau} \underset{T \to \infty}{\to} \frac{c(t)}{1-\beta} \to c(0) = (1-\beta)k(0)
\end{align*}
Thus, optimal consumption path is:
\begin{align*}
c(t+1) = \beta c(t), c(0) = (1-\beta)k(0)
\end{align*}
capital:
\begin{align*}
k(t+1) = k(t) - c(t)
\end{align*}
and the value function is:
\begin{align*}
\sum_{t=0}^{\infty} \beta^t \ln c(t) = \sum_{t=0}^{\infty} \beta^{t} \ln (\beta^t c(0)) = \sum_{t=0}^{\infty} \beta^t t \ln \beta + \sum_{t=0}^{\infty} \beta^t \ln c(0) = \frac{\beta \ln \beta}{(1-\beta)^2} + \frac{\ln c(0)}{1 - \beta } = \\
=\frac{\beta \ln \beta}{(1-\beta)^2} + \frac{\ln((1-\beta)k(0))}{1-\beta}
\end{align*}
\end{enumerate}

\section*{Problem 2}
Problem:
\begin{align*}
\underset{\left\{k(t+1), c(t)\right\}_{t=1}^{\infty}}{\max}\ \sum_{t=1}^{\infty} 0.96^t \ln c(t)\\
\text{s.t.\ } k(t+1) = Ak(t)^{1/3} + (1-0.02)k(t) - c(t)\\
c(t) \ge 0, k(t) \ge 0, k(1) = 0.1
\end{align*}
Euler's equation is:
\begin{align*}
\frac{c(t+1)}{\beta c(t)} = \frac{A}{3k(t+1)^{2/3}} + 1 - \delta
\end{align*}
In the steady state $c(t+1) = c(t)$ hence:
\begin{align*}
\frac{1}{\beta} = \frac{A}{3(k^*)^{2/3}} + 1 - \delta\\
\text{if } k^* = 1 \to A = 3(\frac{1}{\beta} + \delta - 1) = 0.185
\end{align*}
Our initial guess for $k(2)$ should lie within the interval $ [k(1), f(k(1)) + (1 - \delta)k(1)]$ because even if we assume $c(1) = 0$ we will not be able to increase capital $k(2)$ more than $f(k(1)) + (1 - \delta)k(1)$ and moreover since $k^* = 1$ while $k(1) = 0.1$ by the theorem optimal capital growth $s(k(t))$ is non-decreasing in $k(t)$ hence $k(2)$ cannot be less than $k(1)$. The global result which follows is that $k(t)$ cannot be greater than $k^*$ and less than $k(1)$.

Applying shouting algoritm the following value of $k(2)$ has been obtained:
\begin{align*}
k(2) = 0.1368665097187844
\end{align*}
and corresponding capital and consumption paths are depicted on the Fig. \ref{fig1} , \ref{fig2} respectively.
\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\textwidth]{k(t)}
	\caption{}\label{fig1}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{c(t)}
	\caption{}\label{fig2}
\end{figure}

It should be borne in mind that aforementioned $k(2)$ is the $\textit{approximate}$ value of true $k(2)$ that means that with this particular (in general not optimal) $k(2)$ for large $t$ the algorithm will diverge from the steady state. Hence those plots describe true capital and consumption paths only approximately. 
\end{document}