### 2019-01-31 metrics3.tex
Metrics3

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\newcommand{\yrseduc}{\textit{yrseduc}}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Econometrics 1. Problem Set 3.
\end{center}
\section*{Problem 1}

\textbf{Solution}

\begin{enumerate}[(a)]
	\item 
	\begin{align*}
	t = \frac{-0.0023}{0.0022} \approx -1
	\end{align*}
	p-value is about $0.2$, consequently we have no reasons for rejecting the hypothesis that \textit{profitmarg} does not affect salary (i.e. slope coefficient on this variable is equal to zero) on the 5\% significance level.
	\item Similarly,
	\begin{align*}
	t_{2\text{nd col}} = 2.24,\ t_{3\text{nd col}} = 2.04
	\end{align*}
	in both cases the hypothesis that corresponding coeeficient is equal to zero should be rejected on the 5\% significance level. Since
	\begin{align*}
	\frac{\Delta \text{salary}}{\text{salary}} = 0.1 \cdot \frac{\Delta \text{mktval}}{\text{mktval}}\\
	100 \cdot \frac{\Delta \text{salary}}{\text{salary}} = 0.1 \frac{\Delta \text{mktval}}{\text{mktval}} \cdot 100
	\end{align*}
	that means that once \textit{mktval} changes by 1\% holding other thing equal, the average salary will change by 0.1\%.
	\item 
	\begin{align*}
	t_{\text{ceoten}} = \frac{0.0171}{0.0055} = 3.11\\
	\end{align*}
	hence on the 5\% significance level the coefficient is not equal to 0.
	Since
	\begin{align*}
		\frac{\Delta \text{salary}}{\text{salary}} = 0.0171 \Delta \text{ceoten}\\
		100 \cdot \frac{\Delta \text{salary}}{\text{salary}} = 1.71 \Delta \text{ceoten}
	\end{align*}
	that is one more year as a CEO with the current company increases the average salary by 1.71\% other thing equal.
	\begin{align*}
	t_{\text{comten}} = -\frac{0.0092}{0.0033} = 2.78\\
	\end{align*}
	hence on the 5\% significance level the coefficient is not equal to 0.
	Since
	\begin{align*}
	\frac{\Delta \text{salary}}{\text{salary}} = -0.0092 \Delta \text{ceoten}\\
	100 \cdot \frac{\Delta \text{salary}}{\text{salary}} = 0.92 \Delta \text{ceoten}
	\end{align*}
	that is one more year with the current company decreases the average salary by 0.92\% other thing equal.
\end{enumerate}

\section*{Problem 2}
\begin{enumerate}[(a)]
	\item \begin{align*}
	&y_i = \beta_0 + \beta_1 x_{i} + u\\
	&y_i = \gamma_0 + \gamma_1(x_i - \bar{x}) + v = \gamma_0 + \gamma_1 \tilde{x_i} + v\\
	&t_{\beta_1} = \frac{\hat{\beta_1}}{SE(\hat{\beta_1})},\  t_{\gamma_1} = \frac{\hat{\gamma_1}}{SE(\hat{\gamma_1})}
	\end{align*}
	\begin{align*}
	\hat{\beta_1} = \frac{\summa (x_i - \bar{x})y_i}{\summa (x_i - \bar{x})^2} = \frac{\summa \tilde{x_i} y_i}{\summa \tilde{x_i}^2} = \hat{\gamma_1}
	\end{align*}
	\begin{align*}
	SE(\hat{\beta_1}) = \frac{\sqrt{\summa \hat{u_i}^2}}{\sqrt{\summa (x_i - \bar{x})^2}}\\
	SE(\hat{\gamma_1}) = \frac{\sqrt{\summa \hat{v_i}^2}}{\sqrt{\summa (\tilde{x_i})^2}}\\
	\end{align*}
	Thus, the only difference in both standard errors is the difference in sum of squares of residuals. To that end in order to answer the question one need to compare these sums.
	\begin{align*}
	&\hat{u_i} = y_i - \hat{\beta_0} - \hat{\beta_1}x_{i}\\
	&\hat{v_i} = y_i - \hat{\gamma_0} - \hat{\gamma_1}\tilde{x_i} = y_i - \hat{\gamma_0} - \hat{\gamma_1}(x_i - \bar{x})
	\end{align*}
	Since $\hat{\beta_1} = \hat{\gamma_1}$, $\hat{u_i}$ and $\hat{v_i}$ differs by the term $\hat{\gamma_0} - \hat{\gamma_1}\bar{x} - \hat{\beta_0}$.
	\begin{align*}
	&\hat{\gamma_0} = \bar{y} - \hat{\gamma_1}\bar{\tilde{x}} = \bar{y}\\
	&\hat{\beta_0} = \bar{y} - \hat{\beta_1}\bar{x}\\
	&\hat{u_i} - \hat{v_i} = \hat{\beta_1}\bar{x} - \hat{\gamma_1}\bar{x} = 0
	\end{align*}
	Thus, t-statistics are equal.
	\item
	\begin{align*}
	Var \begin{pmatrix}
	\hat{\beta_0}\\
	\hat{\beta_1}
	\end{pmatrix} = \begin{pmatrix}
	2 & 1\\
	1 & 3
	\end{pmatrix}
	\end{align*}
	Since
	\begin{align*}
	\hat{\beta_1} = \beta_1 + \frac{\summa (x_i - \bar{x})u_i}{\summa (x_i - \bar{x})^2}\\
	\end{align*}
	\begin{align*}
	Var (\hat{\beta_1}|x_1, \dots, x_n) \underset{\text{i.i.d}}{=} \frac{\sigma^2_u}{\summa (x_i - \bar{x})^2}
	\end{align*}
	\begin{align*}
	&Var(\hat{\beta_0}|x_1, \dots, x_n) = Var(\bar{y} - \hat{\beta_1}\bar{x}|x_1, \dots, x_n) = Var(\bar{y}|x_1, \dots, x_n) + \bar{x}^2Var(\hat{\beta_1}|x_1, \dots, x_n) -\\
	&- 2\bar{x}cov(\bar{y}, \hat{\beta_1}|x_1, \dots, x_n)
	\end{align*}
	\begin{align*}
	&cov(\bar{y}, \hat{\beta_1}|x_1, \dots, x_n) = \frac{1}{n\summa (x_i - \bar{x})^2} cov \left(\summa y_i, \sum_{j=1}^n (x_j - \bar{x})y_j\bigg|x_1, \dots, x_n\right) = \\
	&=\frac{1}{n\summa (x_i - \bar{x})^2} \sum_{j=1}^n(x_j - \bar{x})\summa cov(y_j, y_i) = \frac{\sigma^2_u \summa (x_i - \bar{x})}{n\summa(x_i - \bar{x})} = 0
	\end{align*}
	Thus, 
	\begin{align*}
	Var(\hat{\beta_0}|x_1, \dots, x_n) = \frac{\sigma^2_u}{n} + \frac{\bar{x}^2\sigma^2_u}{\summa(x_i - \bar{x})^2} = \sigma^2_u\frac{n\bar{x}^2 + \summa(x_i - \bar{x})^2}{n\summa(x_i - \bar{x})^2} = \frac{\summa x_i^2}{n\summa(x_i - \bar{x})^2}\sigma_u^2
	\end{align*}
	\begin{align*}
	cov(\hat{\beta_0}, \hat{\beta_1}|x_1, \dots, x_n) = cov(\bar{y} - \hat{\beta_1}\bar{x}, \hat{\beta_1}|x_1, \dots, x_n) = 0 - \bar{x}Var(\hat{\beta_1}|x_1, \dots, x_n) =\\ -\frac{\bar{x} \sigma^2_u}{\summa (x_i - \bar{x})^2}
	\end{align*}
	To obtain the estimates of variances one should plug $\hat{\sigma_u}^2$ instead of $\sigma^2_u$. That means that:
	\begin{align*}
	\begin{cases}
	\frac{\hat{\sigma^2_u}}{\summa (x_i - \bar{x})^2} = 2\\
	\frac{\summa x_i^2}{n\summa(x_i - \bar{x})^2}\hat{\sigma_u^2} = 3\\
	-\frac{\bar{x} \sigma^2_u}{\summa (x_i - \bar{x})^2} = 1
	\end{cases}\\
	\end{align*}
	Using the fact $\summa (x_i - \bar{x})^2 = \summa x_i^2 - n\bar{x}^2$ we get a system of three equations with three unknows, which can be easily solved, as a result:
	\begin{align*}
	\hat{\sigma^2_u} = 255\\
	\summa x_i^2 = 153\\
	\bar{x} = -\frac{1}{2}
	\end{align*}
	\item Assuming all variances, condition on $x_1, \dots, x_n$
	\begin{align*}
	Var(\hat{\beta_0} + \alpha \hat{\beta_1}) = Var\hat{\beta_0} + \alpha Var \hat{\beta_1} -2\alpha \text{cov}(\hat{\beta_0}, \hat{\beta_1}) =\\
	=\frac{153}{13005}\hat{\sigma_u^2} + \frac{\alpha^2 \hat{\sigma^2_u}}{127.5} + 2\alpha \frac{ \sigma^2_u}{255} \to \min
	\end{align*}
	\begin{align*}
	FOC: \frac{4}{255} \alpha^* + \frac{2}{255} = 0\\
	\alpha^* = -\frac{1}{2}
	\end{align*}
\end{enumerate}
\section*{Problem 3}
The result of regression is depicted below on the Fig. \ref{fig1}.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{lnahe}
	\caption{Summary of regression}\label{fig1}
\end{figure}

\begin{enumerate}[(a)]
	\item First of all, all three coefficients are significant (on 5\% significance level). Of course it would be more correct to use coeftest() in order to get heteroskedasticity-robust standard errors, but p-value is such a small that this procedure do not change our conclusion about significance (only standard errors and t-statistics change slightly). Then, using logic similar to the Problem 1, we can infer that, other things equal each additional year of education increases the average per hour salary by 8\%, holding other things equal each additional year of life increases per hour earnings by 0.6\%, and finally, holding other things equal women get on average 22\% less per hour salary than men.
	\item RMSE is approximately equal to $0.465$.
	\item Confidence interval for the gender gap in earnings is
	\begin{align*}
	(-0.254035195; -0.194271848)
	\end{align*}
	i.e. with probability 0.95 the interval $(19\%; 25\%)$ cover the value which indicates by how much percent the average per hour salary for women is less than the one for men.
	\item Based on the summary of regression, the marginal value of the high school diploma is 8.3, i.e. holding gender and age fixed, receiving a high school diploma after 12 years of studying against 12 years without diploma increases average per hour earnings by 8.3\%. Using coeftest() we can conclude that on the 5\% significance level we reject the hypothesis that the marginal value of high school diploma is zero. 8.3\% is not a huge value, in fact, I suppose average the salary of people with high school diploma much more higher than those without diploma.
	\item 
	Estimating the following regression model
	\begin{align*}
	\ln(\text{ahe}) = \beta_0 + \beta_1 \text{hsdipl} + \beta_2 \text{yrseduc} + \beta_3 \text{age} + \beta_4\text{female} + u
	\end{align*}
	one can get:
	\begin{align*}
	\ln \text{ahe}_{12} - \ln \text{ahe}_{10} &= \beta_1 + 2 \beta_2\\
	\frac{\text{ahe}_{12} - \text{ahe}_{10}}{\text{ahe}_{10}} &= V = e^{\beta_1 + 2\beta_2} - 1 \\
	\hat{V} = e^{\hat{\beta_1} + 2\hat{\beta_2}} - 1 &\approx 0.285
\end{align*}
I think it is a rather reasonable estimate.
\item 
\begin{align*}
\prob (\alpha_1 < e^{\hat{\beta_1} + 2 \hat{\beta_2}} - 1 < \alpha_{2}) = 0.95\\
\prob (\ln(\alpha_1 + 1) < \hat{\beta_1} + 2\hat{\beta_2} < \ln(\alpha_2 +1)) = 0.95
\end{align*}
Thus, we need to compute a confidence interval for $\hat{\beta_1} + 2 \hat{\beta_2}$. To do so let us modify our model and estimate regression $\ln( \text{ahe})$ on \textit{hsdipl } $- 1$, \textit{yrseduc } $-2$, \textit{age} and \textit{female}. It is easy to observe that in that case the intercept will be $\beta_0 + \beta_1 + 2\beta_2$. Using confint() for this new intercept we can get the following:
\begin{align*}
\prob (1.375500454 &< \widehat{\beta_0 + \beta_1 + 2\beta_2} < 1.585969536) = 0.95\\
\prob(0.1458 &< \hat{\beta_1} + 2\hat{\beta_2} < 0.3563) = 0.95
\end{align*}
Thus, $\alpha_1 = 0.157$, $\alpha_2 = 0.428$ and
\begin{align*}
\prob(0.157 < \hat{V} < 0.428) = 0.95
\end{align*}
\end{enumerate}
\section*{Problem 4}
The summary of regression is depicted below (see Fig. \ref{fig2})

\begin{figure}[h]
	\centering
	\includegraphics[width=0.8\textwidth]{square}
	\caption{Summary of the regression}\label{fig2}
\end{figure}

\begin{enumerate}[(a)]
	\item Sign of the coefficient on \textit{age} is +, while on \textit{age}$^2$ is -. Denote $y = \ln \text{ahe}, x_1 = \text{ age}, x_3 = \text{ female}, x_4 = \text{hsdipl}$. Since 
	\begin{align*}
	\ln y = \beta_0 + \beta_1x_1 + \beta_2x_1^2 + \beta_3x_3 + \beta_4x_4 +u
	\end{align*}
	that means that
	\begin{align*}
	\frac{d \ln y}{d x_1} = \beta_1 + 2\beta_2x_1\\
	\frac{dy}{y} = \beta_1dx_1 + 2\beta_2 x_1 dx_1 \\
	\frac{\Delta y}{y} \approx \beta_1\Delta x_1 + 2\beta_2 x_1 \Delta x_1
	\end{align*}
	Since $\beta_1 > 0$ while $\beta_2 < 0$ that means that: the older the person becomes - the less (on average) increase in salary each additional year brings. For example, if $x_1 = 20, \Delta x_1 = 1$ then $\frac{\Delta y}{y} \cdot 100 \approx 2.96 \%$, but if $x_1 = 40, \Delta x_1 = 1$ then $\frac{\Delta y}{y} \cdot 100 \approx 0.86 \%$.
	\item The relation is linear $\iff$ $\beta_2 = 0$. Using coeftest() we reject this hypothesis on the 5\% significance level, i.e. the relation is quadratic.
	\item The estimated gender gap has changed from -0.224 to -0.225, which is not a significant change in a real-world sense. Econometric reason is that first of all $\text{age}^2$ and $\text{female}$ are almost uncorrelated (sample correaltion coeeficient is about 0.007) and moreover slope coefficient on $\text{age}^2$ is also very small, as a result including $\text{age}^2$ barely changes the estimated gender gap.
	\item Let us estimate the following regression
	\begin{align*}
	\ln y = \beta_0 + \beta_1x_1 + \beta_2x_1^2 + \beta_{23}x_1^3+ \beta_3x_3 + \beta_4x_4 +u
	\end{align*}
	the relation is quadratic $\iff \beta_{23} = 0$. Using coeftest() we can conclude that on the 5\% significance level we cannot reject this hypothesis, i.e. the relation is quadratic.
	\item Here we need to test the hypothesis $\beta_2 = 0, \beta_{23} = 0$. The output of linearHypothesis() is depicted below (see Fig. \ref{fig3}).

	\begin{figure}[h]
		\centering
		\includegraphics[width=0.8\textwidth]{hyptest}
		\caption{Hypothesis test}\label{fig3}
	\end{figure}

	Thus, the hypothesis can be rejected on the 5\% significance level. That is the relation is more likely a polynomial of up to cubic degree. I recomend the specification
	\begin{align*}
	\ln y = \beta_0 + \beta_1x_1 + \beta_2x_1^2 + \beta_3x_3 + \beta_4x_4 +u
	\end{align*}
	first of all because this model is more interpretable than the model with cubic relationship, and secondly, we have significant statistic evidence that the relation is not cubic and not linear (based on (d) and (e)).
	\end{enumerate}

\section*{Problem 5}
\begin{enumerate}[(a)]
	\item 
\begin{table}[h]
	\begin{tabular}{cccc}
		Specification & Age 30 to 31  & Age 45 to 46 & Age 60 to 61 \\
		\hline
		Linear in \textit{age}&  &  &  \\
		Quadratic in \textit{age}  &  &  &  \\
		Cubic in \textit{age} &  &  &\\
		\hline 
	\end{tabular}
\end{table}

For linear model the starting point does not matter hence all entries should be equal to each other and from the Problem 3 (a):
\begin{align*}
\frac{\Delta y}{y} \cdot 100 \% \approx 0.6 \Delta x_1
\end{align*}
For quadratic model, as it has been already mentioned above:
\begin{align*}
\frac{\Delta y}{y} &\approx \beta_1 \Delta x_1 + 2 \beta_2x_1\Delta x_1\\
\frac{\Delta y}{y} \cdot 100 \% &\approx 5.07 \Delta x_1 -0.105 x_1 \Delta x_1 = \begin{cases}
1.92, x_1 = 30\\
0.34, x_1 = 45\\
-1.23, x_1 = 60
\end{cases}
\end{align*}
Similarly for cubic model:
\begin{align*}
\frac{\Delta y}{y} \approx \beta_1 \Delta x_1 + 2 \beta_2 x_1 \Delta x_1 + 3\beta_{23}x_1^2 \Delta x_1
\end{align*}
Using summary from the cubic regression we can obtain:
\begin{align*}
\frac{\Delta y}{y} \cdot 100 \% \approx 7.754 \Delta x_1 - 0.234 x_1 \Delta x_1 + 0.0015 x_1^2 \Delta x_1 = \begin{cases}
2.08, x_1 = 30\\
0.26, x_1 = 45\\
-0.886, x_1 = 60
\end{cases}
\end{align*}

Thus, the table looks like:

\begin{table}[H]
	\begin{tabular}{cccc}
		Specification & Age 30 to 31  & Age 45 to 46 & Age 60 to 61 \\
		\hline
		Linear in \textit{age}& 0.6  & 0.6 & 0.6 \\
		Quadratic in \textit{age}  &1.92  &0.34  &-1.23  \\
		Cubic in \textit{age} &2.08  &0.26  &-0.886\\
		\hline 
	\end{tabular}
\end{table}
\item From the linear to quadratic specification this values significantly differs, quadratic specification reflects the truth of life that the average salary firstly increases with age, then it attains maximum at middle age and then starts falling down.

AS for comparison between quadratic and cubic models, both models predict approximately equal values in real world sense.
\item As I have mentioned somewhere above, the second row makes more sense for me, because it is interpretable (formula for percent changes in $y$ does not include $\text{age}^2$ which is a very strange indicator in real-world terms), in contrast to the cubic model and moreover, estimates gotten from this model seem rather plausible.
\end{enumerate}
\end{document}