### 2019-03-16 micro4_2.tex
Micro

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Microeconomics 4. Problem Set 1.
\end{center}
\section*{1\ The Doctor and the Patient}
Consider the cheap talk signalling model a-la Crawford and Sobel (1982). Suppose that
there is a patient that does not feel well and goes to see a doctor. The doctor should
decide which treatment $a$ the patient needs. The type of treatment will depend on
the health state of the patient $\theta$, which is only privately observed by the patient. The
doctor's prior about the patient's health is $\theta \sim U[0; 1]$. The patient can send a costless,
non-binding and non-verifiable signal $s \in [0; 1]$ about his health state (i.e. explain how
he feels). The doctor cares about her reputation, so she wants to prescribe a appropriate
treatment for the patient's health: $U
D(a; \theta) = -(a-\theta)^2$. Instead, the patient, that is
very hypochondriac, wants to be overtreated: $U^P(a; \theta) = -(a - (\theta + b))^2$, where $b > 0$ is
the bias towards excessive medication.
\begin{enumerate}
	\item Assume that b is such that there exist a 2-partitional equilibrium $(p = 2)$. Characterize this equilibrium.
	\item Under which condition will such an equilibrium exist?
	\item Is this the only equilibrium that exists? If yes, argue why. If no, characterize the
	other(s) equilibrium (equilibria).
	\item Compute the expected payoff of the doctor and the patient. How do the payoffs
	change with $b$?
\end{enumerate}

\textbf{Solution}

\begin{enumerate}
	\item Assume that for $\theta < \theta_1$ the patient sends $s_1$ e.g. $s_1 = 0$, while for $\theta \ge \theta_1$ the patient sends $s_2 > s_1$ e.g. $s_2 = \theta_1$. Then the doctor chooses treatment as a maximizer of:
	\begin{align*}
	\begin{cases}
	\underset{a}{\max}\ \int_{0}^{\theta_1} -(a - \theta)^2d\theta, s < \theta_1\\
	\underset{a}{\max}\ \int_{\theta_1}^{1} -(a - \theta)^2d\theta, s \ge \theta_1\\
	\end{cases}
	\end{align*}
	and the solution is obviously
	\begin{align*}
	a(s) = \begin{cases}
	\frac{\theta_1}{2}, s < \theta_1\\
	\frac{1 + \theta_1}{2}, s \ge \theta_1
	\end{cases}
	\end{align*}
	The patient should send:
	\begin{align*}
	\begin{cases}s_1, -\left(\frac{\theta_1}{2} - \theta - b\right)^2 > -\left(\frac{\theta_1 + 1}{2} - \theta - b\right)^2\\
	s_2, \text{ otherwise}
	\end{cases}
	\end{align*}
	That means that $\theta_1$ should be such that the patient becomes indifferent between sending $s_1$ and $s_2$ i.e.
	\begin{align*}
	\frac{\frac{\theta_1}{2} - b + \frac{\theta_1+1}{2} - b}{2} &= \theta_1\\
	\theta_1 &= \frac{1}{2} - 2b
	\end{align*}
	\item 
	\begin{align*}
	\frac{1}{2} - 2b \ge 0\\
	b \le \frac{1}{4}
	\end{align*}
	\item No, there always exists a "babbling" equilibrium. In which the patient always sends the same, uninformative signal and doctor just chooses action which maximize his expected payoff taking in account doctor's prior beliefs about type. Given that, patient indifferent between any signal (he cannot change anything) hence he has not profitable deviation.
	\item The expected payoff of the doctor is:
	\begin{align*}
	\pi^D = \int_{0}^{\frac{1}{2} - 2b} - \left(\frac{1}{4} - b - \theta\right)^2d\theta + \int_{\frac{1}{2} - 2b}^{1}- \left(\frac{3}{4} - b - \theta\right)^2d\theta = -2\frac{\left(\frac{1}{4} - b\right)^3}{3} -2\frac{\left(\frac{1}{4} + b\right)^3}{3}
	\end{align*}
	The expected payoff of the patient is:
	\begin{align*}
	\pi^P = \int_{0}^{\frac{1}{2} - 2b} - \left(\frac{1}{4} - 2b - \theta\right)^2d \theta + \int_{\frac{1}{2} - 2b}^1 - \left(\frac{3}{4} - 2b - \theta\right)^2d \theta = -\frac{1}{3 \cdot 2^5} + \\
	+\frac{1}{3} \left(\left(\frac{3}{4} - 2b\right)^3 + \left(\frac{1}{4} - 2b\right)\right)
	\end{align*}
	\begin{align*}
	&\frac{\partial \pi^D}{\partial b} = 2\left(\frac{1}{4} - b - \frac{1}{4} - b\right)\left(\frac{1}{4} - b + \frac{1}{4} + b\right) = -2b < 0\\
	&\frac{\partial \pi^P}{\partial b} = - 2 \left(\frac{3}{4} - 2b - \frac{1}{4} + 2b\right)\left(\frac{3}{4} - 2b + \frac{1}{4} - 2b\right) = -1 + 4b < 0
	\end{align*}
	That is payoffs goes down once $b$ increases.
\end{enumerate}
\section*{ 2 Labour Market}
There are two types of workers at the labor market, $i = 1, 2$. A worker of type $i$ has a utility function
\begin{align*}
u_i(w, x) = w - c_i(x)
\end{align*}
where $w$ is the wage paid to the worker, $x$ is the output he produces, and $c_i(x)$ is a convex
and increasing cost function, $c'_1(x) < c'_2(x)$, $c_i(0) = 0$, $c'_i(0) = 0$, $c'_i(\infty) = \infty$ for $i = 1, 2$.
The share of workers of the first type is given by $\lambda$.

\begin{enumerate}
	\item Assume that there is a single hiring firm at the market, and it does not observe the
	workers' types. The profit of the firm is given by the value of output less the wage
	payment
	\begin{align*}
	\pi = x - w
	\end{align*}
	Assume that the reservation wage of both workers is zero. Characterize the monopolistic screening contracts $(w_i, x_i)$ offered by the firm.
	\item Now assume that there are multiple firms in the market, but none of them observes
	the workers' types. The firms are competing in hiring workers. Characterize the
	competitive screening SPNE.
	\item How would your answers to (1) and (2) change if the firms can observe the types
	of the workers? Why can the first best be (or not be) achieved in each of these
	scenarios?
	\item Assume that $c_i(x) = \frac{i}{2}x^2,\ i = 1,\ 2$ and $\lambda = 1/2$. Derive the levels of output and the
	wages in your monopolistic and competitive screening problems with unobservable
	and observable worker types (i.e., in (1)-(3))
\end{enumerate}

\textbf{Solution}

\begin{enumerate}
	\item The monopoly's problem is:
	\begin{align*}
	\underset{(x_1, w_1), (x_2, w_2)}{\max}\ \lambda(x_1 - w_1) + (1 - \lambda)(x_2 - w_2)
	\end{align*}
	Since outside options of workers are 0, the optimal contracts monopoly proposes to workers should be such that
	\begin{align*}
	w_1^* = c_1(x_1^*)\\
	w_2^* = c_2(x_2^*)
	\end{align*}
	Assuming interior solution FOC is:
	\begin{align*}
	\begin{cases}
	1 - c_1'(x_1^*) = 0\\
	1 - c_2'(x_2^*) = 0
	\end{cases}\\
	\end{align*}
	\begin{align*}
	c_1'(x_1^*) = c_2'(x_2^*) \to x_1^* > x_2^* \to c_1(x_1^*) > c_1(x_2^*) \to w_2^* - c_1(x_2^*) > w_1^* - c_1(x_1^*) = 0
	\end{align*}
	That means that at any interior solution high type has incentive to deviate and mimic to low type. Hence we should add incentive compatibility constraints. The IC constraints are:
	\begin{align*}
	\begin{cases}
	w_1^* - c_2(x_1^*) \le w_2^* - c_2(x_2^*)\\
	w_2^* - c_1(x_2^*) \le w_1^* - c_1(x_1^*)
	\end{cases}\\
	\begin{cases}
	c_1(x_1^*) - c_2(x_1^*) \le 0\\
	c_2(x_2^*) - c_1(x_2^*) \le 0
	\end{cases}
	\end{align*}
	Since $c_2(x) > c_1(x)\ \forall\ x > 0$ ($g(x) = c_2(x) - c_1(x)$, $g(0) = 0, g'(x) = c'_2(x) - c_1'(x) > 0$ hence $g(x)$ is strictly increasing, since $g(0) = 0$ then $g(x)>0\ \forall\ x > 0$) it follows that $x_2^* = 0$. That means that monopoly should offer the following contracts:
	\begin{align*}
	&(0, 0)\ \text{ which is supposed to be for low type}\\
	&(w_1 = c_1(c_1'^{-1}(1)), x_1 = c_1'^{-1}(1))\ \text{ which is supposed to be for high type}
	\end{align*}
	\item At the competitive equilibrium firms should earn 0 profit. That means that if firms can observe types they should offer the following contracts to both types of workers:
	\begin{align*}
	(x_1, x_1), (x_2, x_2)
	\end{align*}
	Since firms are competing for workers, at the equilibrium both firms should offer workers such contracts that would maximize their utility (otherwise, if any of firm offer another contract, the rival firm can deviate, offer maximizing utility contract, and hire all workers). Of course if firms offer maximizing utility contracts workers have no incentives to deviate and choose a contract which was supposed to be for another type. As a result the firms should propose contracts:
	\begin{align*}
	(c_1'^{-1}(1), c_1'^{-1}(1))\\
	(c_2'^{-1}(1), c_2'^{-1}(1))
	\end{align*}
	\item If monopoly can distinguish between workers, it should propose each type a contract which would maximize its profit and leave each type of workers with zero utility. That is, the contracts should be:
	\begin{align*}
	(w_1 = c_1(c_1'^{-1}(1)), x_1 = c_1'^{-1}(1))\ \text{ high type}\\
	(w_2 = c_2(c_2'^{-1}(1)), x_2 = c_2'^{-1}(1))\ \text{ low type}
	\end{align*}
	If competitive firms can observe types of workers they anyway have to propose such contracts which would maximize workers' utility because otherwise the rival firm hire all workers. That means that firms should propose the same contracts as in the case of unobservable types. That is, first best is achieved under competition and is not achieved under monopoly.
	\item \begin{itemize}
		\item Monopoly, observable types:
		\begin{align*}
		\underset{x_1, x_2}{\max}\ \frac{1}{2}\left(x_1 -\frac{x_1^2}{2}\right) + \frac{1}{2} \left(x_2 - x_2^2\right)\\
		x_1^* = 1,\ x_2^* = \frac{1}{2}\\
		w_1^* = \frac{1}{2},\ w_2^* = \frac{1}{4}
		\end{align*}
		\item Monopoly, unobservable types:
		\begin{align*}
		\underset{x_1, x_2}{\max}\ \frac{1}{2}\left(x_1 -\frac{x_1^2}{2}\right) + \frac{1}{2} \left(x_2 - x_2^2\right)\\
		s.t.\ 	\begin{cases}
		\frac{x_1^2}{2} - x_1^2 \le 0\\
		x_2^2 - \frac{x_2^2}{2} \le 0
		\end{cases}\\
		x_2^* = 0, x_1^* = 1\\
		w_2 = 0, w_1 = \frac{1}{2}
		\end{align*}
		\item Competition, observable types:
		\begin{align*}
		x_1^* = 1, w_1^* = 1\\
		x_2^* = \frac{1}{2}, w_2^* = \frac{1}{2}
		\end{align*}
		\item Competition, unobservable types:
		\begin{align*}
		x_1^* = 1, w_1^* = 1\\
		x_2^* = \frac{1}{2}, w_2^* = \frac{1}{2}
		\end{align*}
	\end{itemize}
\end{enumerate}
\section*{Insurance}
Consider the competitive screening model of the insurance market with two types of
consumers $H$ and $L$. All consumers have initial wealth $W > 0$ and may face a loss $d > 0$,
$d < W$. The probability of facing the loss is $p_L$ for the $L$ type of the consumer, and
$p_H$ for the $H$ type, where $p_H > p_L$. Multiple competitive insurance firms offer insurance
contracts against the loss to either type of the consumer. The probabilities of loss and the
utility functions for both types of consumers are common knowledge to all participants.
For parts (1) and (2), assume that all consumers have the same strictly concave
(Bernoulli) utility of wealth $u(W)$ with $u'(W) > 0$ and $u''(W) < 0$. That is, the two types
of consumers are distinguished only by the difference between $p_L$ and $p_H$
\begin{enumerate}
	\item  Assume that the types of the consumers are publicly observable. Draw a graph
	representing the equilibrium outcome of this model. Explain. Derive the utility levels offered to either type of the consumer in this equilibrium.
	\item Now assume that the types of the consumers are only known to the consumers themselves. Draw a graph to represent the separating equilibrium of this model. Explain the reasoning behind your graph. Compare the outcome to (1) and comment on
	the source of difference/similarity.
	\item Now suppose the $H$ types are more risk averse - with utility function $v(W)$ instead
	of $u(W)$. As above, probabilities of loss and the utility functions for both types of
	consumers are common knowledge, but the types of the consumers are only known
	to the consumers themselves. Draw a new graph to show the effect of this change in
	utility function on the contracts in the separating equilibrium. Suggest an intuitive
	explanation for why this change from $u(W)$ to $v(W)$ increases or reduces the utility
	for good types in the separating equilibrium.
\end{enumerate}

\textbf{Solution}

\begin{enumerate}
	\item 
\end{enumerate}
\end{document}