### 2019-03-23 micro4_2.tex
Micro

\documentclass[a4paper]{article}
\usepackage[14pt]{extsizes} % 
\usepackage[utf8]{inputenc}
\usepackage{setspace,amsmath}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage{titlesec}
\usepackage{pdfpages}
\usepackage[shortlabels]{enumitem}
\usepackage{tikz}
\usetikzlibrary{angles,quotes}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{float}
\usepackage[section]{placeins}
\usepackage[makeroom]{cancel}
\usepackage{mathrsfs} % 
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
%\addto\captionsrussian{\renewcommand{\figurename}{Fig.}}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools} 
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
{\hbox{$\mathsurround=0pt #1$}}{}}
\usepackage{graphicx}  % 
\graphicspath{{images/}{images2/}}  % 
\setlength\fboxsep{3pt} %  \fbox{} 
\setlength\fboxrule{1pt} % \fbox{}
\usepackage{wrapfig} % 
\newcommand{\prob}{\mathbb{P}}
\newcommand{\norma}{\mathscr{N}}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\summa}{\sum_{i=1}^n}
\usepackage[left=7mm, top=20mm, right=15mm, bottom=20mm, nohead, footskip=10mm]{geometry} % 
\usepackage{tikz} % 
\def\myrad{2cm}% radius of the circle
\def\myanga{45}% angle for the arc
\def\myangb{195}
\begin{document} % 
	\begin{flushright}
	\begin{tabular}{r}
		Danil Fedchenko, MAE 2020, group A \\
	\end{tabular}
\end{flushright}

\begin{center}
	Microeconomics 4. Problem Set 1.
\end{center}
\section*{1 The Doctor and the Patient}
Consider the cheap talk signalling model a-la Crawford and Sobel (1982). Suppose that
there is a patient that does not feel well and goes to see a doctor. The doctor should
decide which treatment $a$ the patient needs. The type of treatment will depend on
the health state of the patient $\theta$, which is only privately observed by the patient. The
doctor's prior about the patient's health is $\theta \sim U[0; 1]$. The patient can send a costless,
non-binding and non-verifiable signal $s \in [0; 1]$ about his health state (i.e. explain how
he feels). The doctor cares about her reputation, so she wants to prescribe a appropriate
treatment for the patient's health: $U^D(a; \theta) = -(a-\theta)^2$. Instead, the patient, that is
very hypochondriac, wants to be overtreated: $U^P(a; \theta) = -(a - (\theta + b))^2$, where $b > 0$ is
the bias towards excessive medication.
\begin{enumerate}
	\item Assume that b is such that there exist a 2-partitional equilibrium $(p = 2)$. Characterize this equilibrium.
	\item Under which condition will such an equilibrium exist?
	\item Is this the only equilibrium that exists? If yes, argue why. If no, characterize the
	other(s) equilibrium (equilibria).
	\item Compute the expected payoff of the doctor and the patient. How do the payoffs
	change with $b$?
\end{enumerate}

\textbf{Solution}

\begin{enumerate}
	\item Assume that for $\theta < \theta_1$ the patient sends $s_1$ e.g. $s_1 = 0$, while for $\theta \ge \theta_1$ the patient sends $s_2 > s_1$ e.g. $s_2 = \theta_1$. Then the doctor chooses treatment as a maximizer of:
	\begin{align*}
	\begin{cases}
	\underset{a}{\max}\ \int_{0}^{\theta_1} -(a - \theta)^2d\theta, s < \theta_1\\
	\underset{a}{\max}\ \int_{\theta_1}^{1} -(a - \theta)^2d\theta, s \ge \theta_1\\
	\end{cases}
	\end{align*}
	and the solution is obviously
	\begin{align*}
	a(s) = \begin{cases}
	\frac{\theta_1}{2}, s < \theta_1\\
	\frac{1 + \theta_1}{2}, s \ge \theta_1
	\end{cases}
	\end{align*}
	The patient should send:
	\begin{align*}
	\begin{cases}s_1, -\left(\frac{\theta_1}{2} - \theta - b\right)^2 > -\left(\frac{\theta_1 + 1}{2} - \theta - b\right)^2\\
	s_2, \text{ otherwise}
	\end{cases}
	\end{align*}
	That means that $\theta_1$ should be such that the patient becomes indifferent between sending $s_1$ and $s_2$ i.e.
	\begin{align*}
	\frac{\frac{\theta_1}{2} - b + \frac{\theta_1+1}{2} - b}{2} &= \theta_1\\
	\theta_1 &= \frac{1}{2} - 2b
	\end{align*}
	\item 
	\begin{align*}
	\frac{1}{2} - 2b \ge 0\\
	b \le \frac{1}{4}
	\end{align*}
	\item No, there always exists a "babbling" equilibrium. In which the patient always sends the same, uninformative signal and doctor just chooses action which maximize his expected payoff taking in account doctor's prior beliefs about type. Given that, patient indifferent between any signal (he cannot change anything) hence he has not profitable deviation.
	\item The expected payoff of the doctor is:
	\begin{align*}
	\pi^D = \int_{0}^{\frac{1}{2} - 2b} - \left(\frac{1}{4} - b - \theta\right)^2d\theta + \int_{\frac{1}{2} - 2b}^{1}- \left(\frac{3}{4} - b - \theta\right)^2d\theta = -2\frac{\left(\frac{1}{4} - b\right)^3}{3} -2\frac{\left(\frac{1}{4} + b\right)^3}{3}
	\end{align*}
	The expected payoff of the patient is:
	\begin{align*}
	\pi^P = \int_{0}^{\frac{1}{2} - 2b} - \left(\frac{1}{4} - 2b - \theta\right)^2d \theta + \int_{\frac{1}{2} - 2b}^1 - \left(\frac{3}{4} - 2b - \theta\right)^2d \theta = -\frac{1}{3 \cdot 2^5} + \\
	+\frac{1}{3} \left(\left(\frac{3}{4} - 2b\right)^3 + \left(\frac{1}{4} - 2b\right)\right)
	\end{align*}
	\begin{align*}
	&\frac{\partial \pi^D}{\partial b} = 2\left(\frac{1}{4} - b - \frac{1}{4} - b\right)\left(\frac{1}{4} - b + \frac{1}{4} + b\right) = -2b < 0\\
	&\frac{\partial \pi^P}{\partial b} = - 2 \left(\frac{3}{4} - 2b - \frac{1}{4} + 2b\right)\left(\frac{3}{4} - 2b + \frac{1}{4} - 2b\right) = -1 + 4b < 0
	\end{align*}
	That is payoffs goes down once $b$ increases.
\end{enumerate}
\section*{ 2 Labour Market}
There are two types of workers at the labor market, $i = 1, 2$. A worker of type $i$ has a utility function
\begin{align*}
u_i(w, x) = w - c_i(x)
\end{align*}
where $w$ is the wage paid to the worker, $x$ is the output he produces, and $c_i(x)$ is a convex
and increasing cost function, $c'_1(x) < c'_2(x)$, $c_i(0) = 0$, $c'_i(0) = 0$, $c'_i(\infty) = \infty$ for $i = 1, 2$.
The share of workers of the first type is given by $\lambda$.

\begin{enumerate}
	\item Assume that there is a single hiring firm at the market, and it does not observe the
	workers' types. The profit of the firm is given by the value of output less the wage
	payment
	\begin{align*}
	\pi = x - w
	\end{align*}
	Assume that the reservation wage of both workers is zero. Characterize the monopolistic screening contracts $(w_i, x_i)$ offered by the firm.
	\item Now assume that there are multiple firms in the market, but none of them observes
	the workers' types. The firms are competing in hiring workers. Characterize the
	competitive screening SPNE.
	\item How would your answers to (1) and (2) change if the firms can observe the types
	of the workers? Why can the first best be (or not be) achieved in each of these
	scenarios?
	\item Assume that $c_i(x) = \frac{i}{2}x^2,\ i = 1,\ 2$ and $\lambda = 1/2$. Derive the levels of output and the
	wages in your monopolistic and competitive screening problems with unobservable
	and observable worker types (i.e., in (1)-(3))
\end{enumerate}

\textbf{Solution}

\begin{enumerate}
	\item The monopoly's problem is:
	\begin{align*}
	\underset{(x_1, w_1), (x_2, w_2)}{\max}\ \lambda(x_1 - w_1) + (1 - \lambda)(x_2 - w_2)
	\end{align*}
	Since outside options of workers are 0, the optimal contracts monopoly proposes to workers in case of complete information should be such that
	\begin{align*}
	w_1^* = c_1(x_1^*)\\
	w_2^* = c_2(x_2^*)
	\end{align*}
	Assuming interior solution FOC is:
	\begin{align*}
	\begin{cases}
	1 - c_1'(x_1^*) = 0\\
	1 - c_2'(x_2^*) = 0
	\end{cases}\\
	\end{align*}
	\begin{align*}
	c_1'(x_1^*) = c_2'(x_2^*) \to x_1^* > x_2^* \to c_1(x_1^*) > c_1(x_2^*) \to w_2^* - c_1(x_2^*) > w_1^* - c_1(x_1^*) = 0
	\end{align*}
	That means that at any interior solution high type has incentive to deviate and mimic to low type. Hence we should add incentive compatibility and participation constraints.
	\begin{align*}
	\begin{cases}
	w_1^* - c_2(x_1^*) \le w_2^* - c_2(x_2^*)\\
	w_2^* - c_1(x_2^*) \le w_1^* - c_1(x_1^*)
	\end{cases}
	\begin{cases}
	w_1^* - c_1(x_1^*) \ge 0\\
	w_2^* - c_2(x_2^*) \ge 0
	\end{cases}
	\end{align*}
	Since $c_2(x) > c_1(x)\ \forall\ x > 0$ ($g(x) = c_2(x) - c_1(x)$, $g(0) = 0, g'(x) = c'_2(x) - c_1'(x) > 0$ hence $g(x)$ is strictly increasing, since $g(0) = 0$ then $g(x)>0\ \forall\ x > 0$) it follows that $w_1^* - c_1(x_1^*) \ge w_2^* - c_1(x_2^*) > w_2^* - c_2(x_2^*) \ge 0$. Thus, $w_1^*-c_1(x_1^*) > 0$ is implied by other constraints. Moreover, if $w_2^* - c_2(x_2^*) > 0$ then one can slightly decrease both $w_1^*, w_2^*$ such that all constraints hold but the objective function increases. Hence $w_2^* = c_2(x_2^*)$. On top of that, if $w_1^* - c_1(x_1^*) > w_2^* - c_1(x_2^*)$ then we can slightly decrease $w_1^*$ such that all constraints hold but the objective function increases. That means that $w_1^* - c_1(x_1^*) = w_2 - c_1(x_2^*)$. Finally, the monopoly optimization problem looks as follows:
	\begin{align*}
	\underset{(x_1, w_1), (x_2, w_2)}{\max}\ \lambda(x_1 - w_1) + (1 - \lambda) (x_2 - w_2)\\
	s.t.\ \begin{cases}
	w_1 - c_1(x_1) = c_2(x_2) - c_1(x_2)\\
	w_1 - c_2(x_1) \le 0\\
	c_2(x_2) = w_2
	\end{cases}
	\end{align*}
	FOCs imply:
	\begin{align*}
	\begin{cases}
	c'_1(x_1^*) = 1\\
	\lambda c_1'(x_2^*) + 1 - \lambda - c_2'(x_2^*) = 0
	\end{cases}
	\end{align*}
	If functions $c_1(\cdot), c_2(\cdot)$ are given it is easy to find optimal contracts from FOCs.
	\item At the competitive equilibrium firms should earn 0 profit. That means that if firms can observe types they should offer the following contracts to both types of workers:
	\begin{align*}
	(x_1, x_1), (x_2, x_2)
	\end{align*}
	Since firms are competing for workers, at the equilibrium both firms should offer workers such contracts that would maximize their utility (otherwise, if any of firm offer another contract, the rival firm can deviate, offer maximizing utility contract, and hire all workers). Of course if firms offer maximizing utility contracts workers have no incentives to deviate and choose a contract which was supposed to be for another type. As a result the firms should propose contracts:
	\begin{align*}
	(c_1'^{-1}(1), c_1'^{-1}(1))\\
	(c_2'^{-1}(1), c_2'^{-1}(1))
	\end{align*}
	\item If monopoly can distinguish between workers, it should propose each type a contract which would maximize its profit and leave each type of workers with zero utility. That is, the contracts should be:
	\begin{align*}
	(w_1 = c_1(c_1'^{-1}(1)), x_1 = c_1'^{-1}(1))\ \text{ high type}\\
	(w_2 = c_2(c_2'^{-1}(1)), x_2 = c_2'^{-1}(1))\ \text{ low type}
	\end{align*}
	If competitive firms can observe types of workers they anyway have to propose such contracts which would maximize workers' utility because otherwise the rival firm hire all workers. That means that firms should propose the same contracts as in the case of unobservable types. That is, first best is achieved under competition and is not achieved under monopoly.
	\item \begin{itemize}
		\item Monopoly, observable types:
		\begin{align*}
		\underset{x_1, x_2}{\max}\ \frac{1}{2}\left(x_1 -\frac{x_1^2}{2}\right) + \frac{1}{2} \left(x_2 - x_2^2\right)\\
		x_1^* = 1,\ x_2^* = \frac{1}{2}\\
		w_1^* = \frac{1}{2},\ w_2^* = \frac{1}{4}
		\end{align*}
		\item Monopoly, unobservable types:
		\begin{align*}
		\underset{x_1, w_1, x_2, w_2}{\max}\ \frac{1}{2}\left(x_1 -w_1\right) + \frac{1}{2} \left(x_2 - x_2^2\right)\\
		s.t.\ 	\begin{cases}
		w_1 - \frac{x_1^2}{2} = \frac{x_2^2}{2}\\
		w_1 - x_1^2 \le 0\\
		w_2 = x_2^2
		\end{cases}\\
		x_2^* = \frac{1}{3}, x_1^* = 1\\
		w_2 = \frac{1}{9}, w_1 = \frac{5}{9}
		\end{align*}
		\item Competition, observable types:
		\begin{align*}
		x_1^* = 1, w_1^* = 1\\
		x_2^* = \frac{1}{2}, w_2^* = \frac{1}{2}
		\end{align*}
		\item Competition, unobservable types:
		\begin{align*}
		x_1^* = 1, w_1^* = 1\\
		x_2^* = \frac{1}{2}, w_2^* = \frac{1}{2}
		\end{align*}
	\end{itemize}
\end{enumerate}
\section*{3 Insurance}
Consider the competitive screening model of the insurance market with two types of
consumers $H$ and $L$. All consumers have initial wealth $W > 0$ and may face a loss $d > 0$,
$d < W$. The probability of facing the loss is $p_L$ for the $L$ type of the consumer, and
$p_H$ for the $H$ type, where $p_H > p_L$. Multiple competitive insurance firms offer insurance
contracts against the loss to either type of the consumer. The probabilities of loss and the
utility functions for both types of consumers are common knowledge to all participants.
For parts (1) and (2), assume that all consumers have the same strictly concave
(Bernoulli) utility of wealth $u(W)$ with $u'(W) > 0$ and $u''(W) < 0$. That is, the two types
of consumers are distinguished only by the difference between $p_L$ and $p_H$
\begin{enumerate}
	\item  Assume that the types of the consumers are publicly observable. Draw a graph
	representing the equilibrium outcome of this model. Explain. Derive the utility levels offered to either type of the consumer in this equilibrium.
	\item Now assume that the types of the consumers are only known to the consumers themselves. Draw a graph to represent the separating equilibrium of this model. Explain the reasoning behind your graph. Compare the outcome to (1) and comment on
	the source of difference/similarity.
	\item Now suppose the $H$ types are more risk averse - with utility function $v(W)$ instead
	of $u(W)$. As above, probabilities of loss and the utility functions for both types of
	consumers are common knowledge, but the types of the consumers are only known
	to the consumers themselves. Draw a new graph to show the effect of this change in
	utility function on the contracts in the separating equilibrium. Suggest an intuitive
	explanation for why this change from $u(W)$ to $v(W)$ increases or reduces the utility
	for good types in the separating equilibrium.
\end{enumerate}

\textbf{Solution}

\begin{enumerate}
	\item Both consumers can happen to be in two states of the world:
	\begin{align*}
	W_g &= W - t\\
	W_b &= W - d + r - t
	\end{align*}
	And they derive the following expected utility function:
	\begin{align*}
	V(W_b, W_g) = p_i U(W_b) + (1 - p_i)U(W_g),\ i \in \{L, H\}
	\end{align*}
	Insurance companies' profit is equal to:
	\begin{align*}
	\pi = t - pr
	\end{align*}
	and should be equal to zero. That means that
	\begin{align*}
	W - W_g - p(W_b + d - W_g) = 0\\
	(1 - p)W_g + pW_b = W - pd
	\end{align*}
	Hence break-even lines for high and low types are given as:
	\begin{align*}
	W_b = -\frac{1-p_H}{p_H}W_g + \frac{W}{p_H} -d\\
	W_b = -\frac{1-p_L}{p_L}W_g + \frac{W}{p_L} -d\\	
	\end{align*}
	At the same time, slopes of indifferent curves of consumers are given by:
	\begin{align*}
	\frac{dW_b}{dW_g} = - \frac{\frac{dV}{dW_g}}{\frac{dV}{dW_b}} = -\frac{1-p_i}{p_i}\frac{U'(W_g)}{U'(W_b)}
	\end{align*}
	In complete information case insurance companies can observe the type, hence equilibrium $(W_g^H, W_b^H), (W_g^L, W_b^L)$ should lie on the break-even lines. In other words,
	\begin{align*}
	\frac{dW_b}{dW_g} = - \frac{1-p_i}{p_i}\frac{U'(W_g)}{U'(W_b)} = -\frac{1 - p_i}{p_i}
	\end{align*}
	hence by properties of the utility function $W_g = W_b$ for both types. Thus, both types get full insurance in complete information case. The figure below illustrates the case.
	\begin{figure}[H]
		\centering
		\includegraphics[width=0.8\textwidth]{plotdraft}
		\caption{}\label{fig1}
	\end{figure}
Since $W_g^H = W_b^H, W_g^L = W_b^L$ that means that both types at the equilibrium derive the following utility levels:
\begin{align*}
p_iU(W_b^i) + (1 - p_i)U(W_b^i) = U(W_b^i) = U(W - p_id),\ i \in \{H, L\}
\end{align*}
\item In case of incomplete information at the equilibrium among other things the incentive compatibility constraints should hold. That means that 
\begin{align*}
p_L U(W - d + r_L - t_L) + (1 - p_L)U(W - t_L) \ge p_L U(W - d + r_H - t_H) + \\
+(1 - p_L)U(W - t_H)\\
p_H U(W - d + r_H - t_H) + (1 - p_H)U(W - t_H) \ge p_H U(W - d + r_L - t_L) + \\
+ (1 - p_H)U(W - t_L)\\
\end{align*}
If we assume that as in complete information case $W_g^H = W_b^H, W_g^L = W_b^L$ then IC imply:
\begin{align*}
U(W^L_g) \ge U(W^H_g)\\
U(W^H_g) \ge U(W^L_g)\\
\end{align*}
but by properties of utility functions this implies that $W^L_g = W^H_g$ which is not a separating equilibrium. Thus, at any separating equilibrium the high type should be offered his complete information contract $(W_g^H, W_g^H)$ while the incentive compatibility constraints imply that:
\begin{align*}
p_HU(W_b^H) + (1 - p_H)U(W_g^H) = p_H U(W_b^L) + (1 - p_H)U(W_g^L)\\
U(W - p_Hd) = p_HU(W_b^L) + (1 -p_H)U(W_g^L)
\end{align*}
This separating equilibrium is illustrated below
	\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\textwidth]{plotdraft}
	\caption{}\label{fig2}
\end{figure}
The high risk type obtains the same utility level as in complete information case, whereas the low risk type becomes strictly worse-off, since $W^g_L \neq W_b^L$. The source of this is the existence of high risk individual who wants to mimic to low risk type and make insurance companies to lose money.
\item If high type is more risk averse, then for a given wealth level in good state he will demand more wealth in bad state comparing with utility $u(W)$ to maintain the same utility level. Put it another way, if less risk averse consumer is indifferent between full insurance and some other point, then more risk averse consumer strictly prefers complete insurance. That means that new separating equilibrium looks like as follows:
	\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\textwidth]{plotdraft}
	\caption{}\label{fig3}
\end{figure}
As we can see, in new equilibrium low type obtain higher level of utility, however still less than in the complete information case. Intuitively it happens because once the high risk type becomes more risk averse, the amount of riskier (with not complete insurance) contracts which attract this type decreases. As a result, now the low type becomes happier, obtaining the contract which is closer to complete insurance. 
\end{enumerate}

\section*{4 Productive Tasks}
	Extend the job market screening model that we saw in Lecture 8 to the case in which
	tasks are productive. Assume that a type $\theta$ worker produces $\theta(1 + \mu t)$ units of output
	when her task level is $t$ where $\mu > 0$.
	Identify the SPNE of this game.

	\textbf{Solution}
	SPNE is depicted below.
	\begin{figure}[H]
		\centering
		\includegraphics[width=0.8\textwidth]{plotdraft}
		\caption{}\label{fig4}
	\end{figure}
	As in usual screening model pooling equilibria are impossible. And separating equilibrium is on the picture. It is indeed equilibria because in this case it cannot be beaten a separating equilibrium as in unproductive task model, because in this equilibrium high type obtains higher level of utility than in any pooling equilibrium. And moreover since indifference curve of high type is tangent to the line $w = \theta_H(1 + \mu t)$ firms cannot simultaneously offer a slightly worse (with negative profit) contract for the low type and slightly better (with positive profit) for low type, because the high type will become strictly worse-off.

	\section*{5 The Patient}
		Suppose that the hypochondriac patient of question 1 is a professional runner. The
		patient has been training for a very important race for the last two years and he is very
		much looking forward to it. In order to run in this important race, first he should be
		cleared that he is "fit for running" by the Organization of the race. Suppose that when
		the patient meets his doctor, there is also a member of the Organization in the surgery.
		What do you think will happen to the quality of the information transmitted by the
		patient?

		\textbf{Solution}

		Since the patient is very much looking forward to taking part in the race, he does not want to be declared unfit for running. That means parameter $b$ effectively becomes smaller (of course parameter $b$ is the innate trait of the agent and cannot change, but once the member of Organization is in the surgery, the patient becomes to obtain disutility from the overtreating and declaring unfit for the race, i.e. his utility becomes $U^P = - (a - (b + \theta - m))^2$ where $m > 0$. If we denote $b' = b - m$ then his effective bias $b'$ becomes smaller once the member of Organization is in surgery). Because if patient complains about his health a lot, the member of the Organization will likely to declare him as unfit for running. And the sportsman understands it. As is has been shown in Problem 1, once bias becomes smaller, expected profits increases. The same result was shown for a more general set-up by Crawford and Sobel (1982):
		\begin{quote}
			There may be a good case for presuming that direct communication is more likely to play an important role, the more closely related are agents' goals.
		\end{quote}
	That is, the quality of information transmission is likely to become higher.

		\section*{6 Absence of Equilibria}
			In the job market screening model that we saw in class, we found that a pooling contract
			was not an equilibrium because a firm could make money by skimming off the good types.
			We also saw that a separating equilibrium is beaten by a pooling contract under some
			conditions. Do you view this result as troubling? How might we be able to alter the
			model to get a more reasonable prediction?

			\textbf{Solution}

			We should use some refinements i.e. slightly alter the concept of equilibrium. We can use anticipatory (or E2) (Wilson, JET, 1977 ) and reactive (Riley, Econometrica, 1979) equilibrium concepts.
			Wilson (1977) defines the anticipatory equilibrium as follows:
			\begin{quote}
				Each policy in the equilibrium set must earn nonnegative profits and there must not be a new set of policies which will earn nonnegative profits for each policy and strictly positive profits for some new policy after the policies in the original set which have become unprofitable are withdrawn
			\end{quote} 
		As a result, this concept manages to deal with absence of pooling equilibria. Once the previous contracts are withdrawn under this concept, firm skim off not only good type but also bad type, but in this situation firms begin losing money, so the deviation is no longer profitable. That is, once contracts are withdrawn firms cannot skim off only a good type.
			Riley (1979) defines the concept of reactive equilibrium as follows:
			\begin{quote}
				A set of offers is a reactive equilibrium if, for any additional offer which generates as expected gain to the agent making the offer, there is another which yields a gain to a second agent and losses to the first. Moreover, no further addition to the set of offers generates losses to the second agent.
			\end{quote}
			This concept manages to deal with the second problem. It allows firms to add more contracts. Then, if we are in the case when separating equilibrium is beaten by pooling, adding another contract (see Fig. \eqref{fig5}, point 2) makes the initial deviation no longer profitable. Hence the separating equilibrium is not beaten by pooling under this concept.
				\begin{figure}[H]
				\centering
				\includegraphics[width=0.8\textwidth]{plotdraft}
				\caption{}\label{fig5}
			\end{figure}

		\section*{7 Single-crossing}
			Most screening models assume single-crossing condition. Think about what this condition
			would require in real life. Do you think single-crossing is always a reasonable assumption?
			Explain why it is reasonable in any application, or suggest a real-life example where this assumption might fail and explain why.

			\textbf{Solution}

			This condition requires that whatever effort people have already put to signal of their type, for putting the additional amount of effort the low type always requires more compensation than the high type. I think that it is a reasonable assumption in real-life terms but only starting from some point. Cost of some signals can be very low, as a result, both groups of people will demand the same amount of compensation for completing these tasks. For example, almost everybody graduates from school and the vast majority then enter in some universities. The costs of this signals are very low, so people of both types of abilities bear the similar costs of signaling. But starting from some point (on effort line), tasks become more complicated and require more time and cognitive abilities, in this case people of high type will spend less time, put less effort to solve such a kind of tasks. For example just to graduate from school is a very easy task, however graduate from school with some kind of honors (gold medal or smth like that) is not so easy, as a result the majority of people who succeed in it are either of high type or of low type but who put a lot of effort (maybe these efforts transformed them to people of high type) and here the single crossing condition is satisfied. I think that true indifference curves should look something like this:
			 \begin{figure}[H]
			 	\centering
			 	\includegraphics[width=0.8\textwidth]{plotdraft}
			 	\caption{}\label{fig6}
			 \end{figure}
\end{document}